---
output: pdf_document
---
# Appendix
## Preprocessing
```{r warning=FALSE}
# Import packages
library(readr)
library(MASS)
library(forecast)
```

```{r}
dt <- read.csv("~/Desktop/pstat174/project/new_cases.csv") # Load data
cp <- dt # Make a copy of data
cp["World"][cp["World"]==0] = 0.00001 # Subtitute 0s for future analyzing
cvnc_row <- nrow(cp) 
cvnc <- cp[2:cvnc_row, 3]
cvnc_ts = ts(cvnc, start = 1)
```

```{r fig.show='hide'}
# Plot orginal time series
ts.plot(cvnc_ts, ylab = "New confirmed cases", main = "Raw Data with Mean and Trend Lines")
abline(a=mean(cvnc_ts), b = 0, col="red") # Mean line
abline(lm(cvnc_ts ~ as.numeric(1:length(cvnc_ts))), col="blue") # Trend line
```

```{r}
# Split data
cvnc_row
cvnc.train <- cvnc_ts[c(1:140)] # Training data
cvnc.test <- cvnc_ts[c(141:152)] # Test data
```

```{r fig.show='hide'}
# Plot training data
ts.plot(cvnc.train, ylab = "New confirmed Cases", main = "Training Data with Mean and Trend Lines")
abline(a=mean(cvnc.train), b = 0, col="red") # Mean line
abline(lm(cvnc.train ~ as.numeric(1:length(cvnc.train))), col="blue") # Trend line
```

```{r fig.show='hide'}
# Plot histogram of training data
par(mfrow=c(1, 2))
hist(cvnc.train, main = "Histogram of Training Data", xlab = "New confirmed cases")
acf(cvnc.train, main = "ACF of Training Data")
```

## Transformation
```{r fig.show='hide'}
# Find lambda for Box-Cox transformation  
t <- 1:length(cvnc.train)
bcTransform <- boxcox(cvnc.train ~ t, plotit = TRUE)
```

```{r}
# Apply Box-Cox transformation
lambda <- bcTransform$x[which(bcTransform$y == max(bcTransform$y))]
lambda
cvnc.train.bc <- (1/lambda)*(cvnc.train**lambda-1)
```

```{r fig.show='hide'}
# Plot transformed training data
par(mfrow=c(1, 2))
ts.plot(cvnc.train, ylab = "New confirmed cases", main = "Original Training Data")
ts.plot(cvnc.train.bc, main = "Transformed Training Data", ylab = expression(Y[t]))
```

```{r fig.show='hide'}
# Plot histogram of transformed data
par(mfrow=c(1, 2))
hist(cvnc.train, main = "Histogram of Original Data")
hist(cvnc.train.bc, main = "Histogram of Transformed Data")
```

## Differencing
```{r fig.show='hide'}
# Differencing at lag = 1 once
cvnc.train1 <- diff(cvnc.train.bc, 1)
# Plot transformed data with differencing once
ts.plot(cvnc.train1, main = "Transformed Training Data with Differencing at Lag = 1 Once")
# Compute variance
var(cvnc.train1)
```

```{r fig.show='hide'}
# Differencing at lag = 1 twice
cvnc.train2 <- diff(cvnc.train1, 1)
# Plot transformed data with differencing twice
ts.plot(cvnc.train2, main = "Training Data with Differencing at Lag = 1 Twice")
# Compute variance
var(cvnc.train2)
```

```{r fig.show='hide'}
# Plot data after transformation and differencing
par(mfrow=c(2, 2))

ts.plot(cvnc.train, ylab = "New confirmed cases", main = "Original Data")
abline(a=mean(cvnc.train), b = 0, col="red") # Mean line
abline(lm(cvnc.train ~ as.numeric(1:length(cvnc.train))), col="blue") # Trend line

ts.plot(cvnc.train.bc, main = "Box-Cox Transformed Data", ylab = expression(Y[t]))
abline(a=mean(cvnc.train.bc), b = 0, col="red") # Mean line
abline(lm(cvnc.train.bc ~ as.numeric(1:length(cvnc.train.bc))), col="blue") # Trend line

ts.plot(cvnc.train1, ylab = "Transformed_1", main = "De-trended Data") 
abline(a=mean(cvnc.train1), b = 0, col="red") # Mean line
abline(lm(cvnc.train1 ~ as.numeric(1:length(cvnc.train1))), col="blue") # Trend line

```

```{r fig.show='hide'}
# Plot acf for data after transformation and differencing
par(mfrow=c(2, 2))
acf(cvnc.train, main = "ACF of Original Data")
acf(cvnc.train.bc, main = "ACF of Transformed Data")
acf(cvnc.train1, main = "ACF of De-trended Data")
```

```{r fig.show='hide'}
# Plot histogram for data after transformation and differencing
par(mfrow=c(2, 2))
hist(cvnc.train, main = "Histogram of Original Data")
hist(cvnc.train.bc, main = "Histogram of Transformed Data")
hist(cvnc.train1, main = "Histogram of De-trended Data")
```

## Model selection
```{r fig.show='hide'}
par(mfrow=c(1, 2))
acf(cvnc.train1, main = "ACF of De-trended Data")
pacf(cvnc.train1, main = "PACF of De-trended Data")
```

ACF outside confidence intervals: lags 1, 2, 8, 11 and 16\newline
PACF outside confidence intervals: lags 1 and 6\newline

```{r}
# Pure moving average models
arima(cvnc.train.bc, order=c(0,1,2), method="ML") 
arima(cvnc.train.bc, order=c(0,1,8), method="ML")
arima(cvnc.train.bc, order=c(0,1,11), method="ML") 
arima(cvnc.train.bc, order=c(0,1,16), method="ML") 
```

```{r}
# Pure autoregressive models 
arima(cvnc.train.bc, order=c(1,1,0), method="ML") 
arima(cvnc.train.bc, order=c(6,1,0), method="ML") 
```

```{r}
# ARIMA models
arima(cvnc.train.bc, order=c(1,1,2), method="ML") 
arima(cvnc.train.bc, order=c(1,1,8), method="ML") 
arima(cvnc.train.bc, order=c(1,1,11), method="ML") 
arima(cvnc.train.bc, order=c(1,1,16), method="ML") 
```

```{r}
arima(cvnc.train.bc, order=c(6,1,2), method="ML") 
arima(cvnc.train.bc, order=c(6,1,8), method="ML") 
arima(cvnc.train.bc, order=c(6,1,11), method="ML") 
arima(cvnc.train.bc, order=c(6,1,16), method="ML") 
```

Model arima(x = cvnc.train.bc, order = c(6, 1, 11), method = "ML") has lowest AIC but the number of significant coefficients are too many to continue future analyze. Therefore, we choose model starting with second lowest AIC.\newline

## Fix models
```{r}
# Selection 1
arima(x = cvnc.train.bc, order = c(6, 1, 16), method = "ML") 
```

```{r}
# Fix coefficients
arima(x = cvnc.train.bc, order = c(5, 1, 16), method = "ML", fixed = c(NA, 0, NA, NA, NA, 0, 0, NA, 0, NA, 0, 0, 0, 0, 0, NA, 0, NA, NA, 0, NA)) 
```


```{r}
# Check stationarity
source("plot.roots.R")
plot.roots(NULL, polyroot(c(1, -0.4807, 0, 0.5175, 0.4361, -0.3167)), main="Autoregressive Part")
```

```{r}
# Selection 2
arima(x = cvnc.train.bc, order = c(1, 1, 16), method = "ML") 
```

```{r}
# Fix coefficients
arima(x = cvnc.train.bc, order = c(1, 1, 16), method = "ML", fixed = c(NA, NA, NA, 0, 0, NA, NA, 0, NA, 0, NA, 0, 0, NA, NA, 0, NA)) 
```

```{r}
# Check invertibility
plot.roots(NULL, polyroot(c(1, 0.3738, -0.3525, 0, 0, 0.3782, 0.3611, 0, 0.1749, 0, -0.0179, 0, 0, 0.1672, 0.1452, 0, 0.2646)), main="Moving Average Part")
```

```{r}
# Selection 3
arima(x = cvnc.train.bc, order = c(1, 1, 11), method = "ML")
```

```{r}
# Fix coefficients
arima(x = cvnc.train.bc, order = c(1, 1, 11), method = "ML", fixed = c(NA, NA, NA, NA, 0, NA, NA, 0, 0, 0, 0, NA))
```

```{r}
# Check invertibility
plot.roots(NULL, polyroot(c(1, 0.5515, -0.3294, 0.0388, 0, 0.4063, 0.4455, 0, 0, 0, 0, -0.1821)), main="Moving Average Part")
```

```{r}
# Selection 4
arima(x = cvnc.train.bc, order = c(6, 1, 8), method = "ML")
```

```{r}
# Fix coefficients
arima(x = cvnc.train.bc, order = c(6, 1, 8), method = "ML", fixed = c(NA, 0, 0, 0, NA, NA, NA, NA, NA, NA, NA, NA, 0, NA))
```

```{r}
# Check stationarity
plot.roots(NULL, polyroot(c(1, 0.7413, 0, 0, 0, -0.5473, 0.5757)), main="Autoregressive Part")
```

```{r}
# Selection 5
arima(cvnc.train.bc, order=c(1,1,8), method="ML")
```

```{r}
# Fix coefficients
arima(cvnc.train.bc, order=c(1,1,6), method="ML", fixed = c(NA, NA, NA, 0, 0, NA, NA))
```

```{r}
# Check invertibility
plot.roots(NULL, polyroot(c(1, 0.5265, -0.3411, 0, 0, 0.4171, 0.4931)), main="Moving Average Part")
```

```{r}
# Selection 6
arima(cvnc.train.bc, order=c(0,1,16), method="ML")
```

```{r}
# Fix coefficients
arima(cvnc.train.bc, order=c(0,1,16), method="ML", fixed = c(NA, 0, 0, 0, NA, 0, 0, NA, 0, 0, 0, NA, 0, NA, 0, NA))
```

```{r}
# Check invertibility 
plot.roots(NULL, polyroot(c(1, -0.8001, 0, 0, 0, 0.3163, 0, 0, 0.2673, 0, 0, 0, 0.3731, 0, 0.2879, 0, 0.1556)), main="Moving Average Part")
```

```{r}
# Selection 7
arima(x = cvnc.train.bc, order = c(0, 1, 8), method = "ML") 
```


```{r}
# Fix coefficients 
arima(x = cvnc.train.bc, order = c(0, 1, 8), method = "ML", fixed = c(NA, NA, 0, 0, NA, 0, 0, NA)) 
```

```{r}
# Check invertibility
plot.roots(NULL, polyroot(c(1, -0.4873, 0.1936 , 0, 0, 0.2386, 0, 0, 0.1819)), main="Moving Average Part")
```

Final model chosen:\newline
Model 1: $(1+0.9575B_{(0.0411)})(1-B)X_t = (1+0.5265_{(0.0842)}B-0.3411_{(0.0692)}B^2+0.4171_{(0.1059)}B^5+0.4931_{(0.1043)}B^6)Z_t$\newline
Model 2: $(1-B)X_t = (1-0.4873_{(0.0768)}B+0.1936_{(0.0903)}B^2+0.2386_{(0.1245)}B^5+0.1819_{(0.0850)}B^8)Z_t$\newline


##Diagnostic checking
```{r fig.show='hide'}
# Model 1
fit1 <- arima(cvnc.train.bc, order=c(1,1,6), method="ML", fixed = c(NA, NA, NA, 0, 0, NA, NA))
res1 <- residuals(fit1)

par(mfrow=c(2, 2))

# Plot histogram of res1
hist(res1, main = "Histogram of Residuals of Model 1")

# Plot res1
ts.plot(res1, main = "Residuals of Model 1")
abline(a=mean(res1), b = 0, col="red") # Mean line
abline(lm(res1 ~ as.numeric(1:length(res1))), col="blue") # Trend line

# Q-Q plot of res1
qqnorm(res1)
qqline(res1, col = "Blue")

# Compute average
mean(res1)
```

```{r fig.show='hide'}
# Plot acf and pacf for res1
par(mfrow=c(1, 2))
acf(res1, main = "ACF of Residuals of Model 1")
pacf(res1, main = "PACF of Residuals of Model 1")
```

```{r}
# Compute approximate value of lag
sqrt(length(cvnc.train))
```

```{r results='hide'}
# Residual tests for residuals of model 1
shapiro.test(res1)
Box.test(res1, lag = 12, type = c("Box-Pierce"), fitdf = 5)
Box.test(res1, lag = 12, type = c("Ljung-Box"), fitdf = 5)
Box.test(res1^2, lag = 12, type = c("Ljung-Box"), fitdf = 0)
ar(res1, aic = TRUE, order.max = NULL, method = c("yule-walker"))
```

```{r fig.show='hide'}
# Model 2
fit2 <- arima(x = cvnc.train.bc, order = c(0, 1, 8), method = "ML", fixed = c(NA, NA, 0, 0, NA, 0, 0, NA)) 
res2 <- residuals(fit2)

par(mfrow=c(2, 2))

# Plot histogram of res2
hist(res2, main = "Histogram of Residuals of Model 2")

# Plot res2
ts.plot(res2, main = "Residuals of Model 2")
abline(a=mean(res2), b = 0, col="red") # Mean line
abline(lm(res2 ~ as.numeric(1:length(res2))), col="blue") # Trend line

# Q-Q plot of res2
qqnorm(res2)
qqline(res2, col = "Blue")

# Compute average
mean(res2)
```

```{r fig.show='hide'}
# Plot acf and pacf for res2
par(mfrow=c(1, 2))
acf(res2, main = "ACF of Residuals of Model 2")
pacf(res2, main = "ACF of Residuals of Model 2")
```

```{r results='hide'}
# Residual tests for residuals of model 2
shapiro.test(res2)
Box.test(res2, lag = 12, type = c("Box-Pierce"), fitdf = 4)
Box.test(res2, lag = 12, type = c("Ljung-Box"), fitdf = 4)
Box.test(res2^2, lag = 12, type = c("Ljung-Box"), fitdf = 0)
ar(res2, aic = TRUE, order.max = NULL, method = c("yule-walker"))
```

## Forecast
```{r}
# Forecast with model 1
forecast(fit1, 13) 
pred.tr <- predict(fit1, n.ahead = 13)

# Compute confidence intervals
U.tr= pred.tr$pred + 2*pred.tr$se
L.tr= pred.tr$pred - 2*pred.tr$se

#Transformation for complete data
cvnc.bc <- (1/lambda)*(cvnc**lambda-1)
```

```{r fig.show='hide'}
# Plot prediction on transformed data
ts.plot(cvnc.train.bc, xlim=c(1,length(cvnc.train.bc)+13), ylim = c(min(cvnc.train.bc), max(U.tr)))
lines(U.tr, col="blue", lty="dashed")
lines(L.tr, col="blue", lty="dashed")
points((length(cvnc.train.bc)+1):(length(cvnc.train.bc)+13), pred.tr$pred, col="red")
points((length(cvnc.train.bc)+1):(length(cvnc.train.bc)+13), cvnc.bc[(length(cvnc.train.bc)+1):(length(cvnc.train.bc)+13)], col="green")
```

```{r}
# Compute prediction without transformation
pred.orig <- (pred.tr$pred*lambda+1)**(1/lambda)
U <- (U.tr*lambda+1)**(1/lambda)
L <- (L.tr*lambda+1)**(1/lambda)
```

```{r fig.show='hide'}
# Plot prediction on original data
ts.plot(cvnc.train, xlim=c(1,length(cvnc.train)+13), ylim = c(min(cvnc.train), max(U)), main = "COVID-19 New Confirmed Cases", ylab = "New confirmed cases")
lines(U, col="blue", lty="dashed")
lines(L, col="blue", lty="dashed")
points((length(cvnc.train)+1):(length(cvnc.train)+13), pred.orig, col="red")
points((length(cvnc.train)+1):(length(cvnc.train)+13), cvnc[(length(cvnc.train)+1):(length(cvnc.train)+13)], col="green")
```

```{r fig.show='hide'}
# Plot prediction on original data
ts.plot(cvnc, ylim = c(min(cvnc.train), max(U)), main = "COVID-19 New Confirmed Cases", ylab = "New confirmed cases")
lines(U, col="blue", lty="dashed")
lines(L, col="blue", lty="dashed")
points((length(cvnc.train)+1):(length(cvnc.train)+13), pred.orig, col="red")
```

```{r fig.show='hide'}
# Plot zoomed prediction on original data
ts.plot(cvnc.train, xlim = c(130, 160), ylim = c(min(L), max(U)), main = "COVID-19 New Confirmed Cases", ylab = "New confirmed cases")
lines(U, col="blue", lty="dashed")
lines(L, col="blue", lty="dashed")
points((length(cvnc.train)+1):(length(cvnc.train)+13), pred.orig, col="red")
points((length(cvnc.train)+1):(length(cvnc.train)+13), cvnc[(length(cvnc.train)+1):(length(cvnc.train)+13)], col="green")
```

```{r fig.show='hide'}
# Plot zoomed prediction on original data
ts.plot(cvnc, xlim = c(130, 160), ylim = c(min(L), max(U)), main = "COVID-19 New Confirmed Cases", ylab = "New confirmed cases")
lines(U, col="blue", lty="dashed")
lines(L, col="blue", lty="dashed")
points((length(cvnc.train)+1):(length(cvnc.train)+13), pred.orig, col="red")
```